Teaching Low-Level Concurrency with \pebbles and \landslide

### 2016-08-23 sigcse-17/landslide.tex
sigcse: shorten some things and lengthen others

{\bf State space exploration.}
To trigger the interleaving which exposes this bug, threads must be preempted at the boundaries of the synchronization primitives (either when \x{cond_wait} blocks or at the end of \x{mutex_unlock}).
Landslide automatically considers these boundaries to be {\em preemption points}: candidate sites for timer-driven thread switches.
It can also add preemption points dynamically,
as identified by an integrated data-race analysis \cite{tsan,fasttrack},
to expose shared-memory bugs when synchronization is absent.

@@

Student submissions may behave in extraordinary ways which violate the assumptions of many modern model checkers.
Landslide refuses to test code which fails the hurdle criteria (\sect{\ref{sec:grading}}),
but it must support testing any legal, yet distasteful, solution.
%Hence, we added several heuristic
For example, students may use open-coded \x{yield} or \x{xchg} loops rather than the mutex abstraction. %(though it is discouraged).
To avoid getting stuck in a cyclic (infinite) state space, Landslide keeps a ``yield loop'' counter for each thread;
when a thread \x{yield}s 10 times without otherwise making progress, Landslide marks it blocked.
%
In extreme cases, this may erroneously seem like deadlock when all threads \x{yield} many times coincidentally.
To avoid wasting student time with false alarms, % the ultimate sin
Landslide will reset all blocked threads' yield-loop counters,
retrying potentially up to 128 times before reporting a true deadlock.

Students may also break abstraction by open-coding a \x{mutex_unlock} sequence.
Landslide must track lock operations for accurate data-race analysis,
so to catch these patterns, it
%learns the size of the \x{mutex} struct, then
checks accesses against stored lock addresses blocked on by other threads.
Similarly, these rogue unlock accesses may unblock a yield-looping thread.

>>>>

{\bf State space exploration.}
To trigger the interleaving which exposes this bug, threads must be preempted at the boundaries of the synchronization primitives (either when \x{cond_wait} blocks or at the end of \x{mutex_unlock}).
Landslide automatically considers these boundaries to be {\em preemption points}: candidate sites for timer-driven thread switches.
It can also add preemption points dynamically,
as identified by an integrated data-race analysis \cite{hybriddatarace,fasttrack},
to expose shared-memory bugs when synchronization is absent.

@@

Student submissions may behave in extraordinary ways which violate the assumptions of many modern model checkers.
Landslide refuses to test code which fails the hurdle criteria (\sect{\ref{sec:grading}}),
but it must support testing any legal, yet distasteful, solution.
%Hence, we added several heuristic
For example, students may use open-coded \x{yield} or \x{xchg} loops rather than the mutex abstraction. %(though it is discouraged).
To avoid getting stuck in a cyclic (infinite) state space, Landslide keeps a ``yield loop'' counter for each thread;
when a thread \x{yield}s 10 times without otherwise making progress, Landslide marks it blocked.
%
In extreme cases, this may erroneously seem like deadlock when all threads \x{yield} many times coincidentally.
To avoid wasting precious student attention with false alarms, % the ultimate sin
Landslide will reset all blocked threads' yield-loop counters,
retrying potentially up to 128 times before reporting a true deadlock.

Students may also break abstraction by open-coding a \x{mutex_unlock} sequence.
Landslide must track lock operations for accurate data-race analysis,
so to catch these patterns, it
%learns the size of the \x{mutex} struct, then
checks accesses against stored lock addresses blocked on by other threads.
Similarly, these rogue unlock accesses may unblock a yield-looping thread.
These heuristics have enabled Landslide to test over 100 student thread libraries from the past 5 semesters fully automatically.
